### 2017-09-14 paper/biblio/security.tex

\documentclass{article}
\usepackage{graphicx,color,setspace,hyperref}
\usepackage{mathtools}
\usepackage{hyperref}
\usepackage{listings}
\lstset{language=C}

\begin{document}

\title{Security}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Motivation
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Motivation}

\noindent\cite{garfinkel:NDSSS2003-VM-introspection} VM IDSs - the paper provides good motivation for the use of virtual machine IDSs and their abilities contribute to security. They prove that deviations from the model and reality can allow evasion.\\

\noindent\cite{laureano:euromicro2004-idsInVMs} is the pioneering work and motivation for virtual machine IDSs. Discusses VMs, hypervisors, \& introspection (monitor \& respond).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introspection}

\noindent\cite{dolan-gavitt:ieeesp2011-virtuoso} Virtuoso - a system that edits QEMU assembly translations, outputs trace logs, analyzes/consolidates the trace logs, \& creates a host python simulation (including IO).\\

\noindent\cite{bahram:IEEErds2012-DKSM} DKSM - a LKM that changes in-kernel data structures (using QEMU + IL) to thwart introspection. \\

\noindent\cite{payne:acsac2007-secure-flexible-vm-monitoring} XenAccess -  provides the foundation introspection tools viw virtual {memory introspection, virtual disk monitoring}.\\

\noindent\cite{fu:sp2012-space-traveling} VMST - extension to Virtuoso the *automatically* bridges the semantic gap to generate VMI tools - (user doesn't specify tools); system wide instruction monitoring to redirect important kernel data. \\

\noindent\cite{dolan-gavitt:techreport2011-leveraging-forensic-introspection} FMA tools can be applied to live VMI-introspection. \\

\noindent\cite{toldinas:kaunas2009-VM-rootkits} memory snapshots - viability and drawbacks. They varied VM memory size, and, in addition to other affects, concluded that VM memory size is correlated to the root kit detection time.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternatives to Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Alternate to introspection}

\noindent\cite{vasisht:micro2008-SHARK} SHARK - proposes micro architectural support for a synergistic hardware/software system to detect root kits - it completely distrusts the OS.\\

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Alternatives to Introspection
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Class Readings}

\noindent\cite{gentry:acm2010-homomorphic} talks about homomorphic encryption schemes which may be helpful for interactions with the cloud.\\

\noindent\cite{rivers:acm1978-onewayf} the fundamental basis for encryption - a function that is easy to compute but difficult to reverse.\\

\noindent\cite{chen:sp2010-side-channels} side channels - how user clicks or keystrokes can reveal information.\\

\noindent\cite{dingledine:ssym2004-tor} Tor - system that pads data and encrypts messages with relays.\\

\bibliographystyle{abbrv}
\bibliography{references}

\addcontentsline{toc}{section}{References}

\newpage
\onecolumn
\appendix
%Appendix A
%\section{1}

\end{document}