### 2017-09-27 paper/sections/6_conclusion.tex
paper: new outline based on SSIO feedback
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Conclusion}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_size.png}\\
  \caption{\label{fig:keyspace-analysis_size}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_locality.png}\\
  \caption{\label{fig:keyspace-analysis_locality}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_throughput.png}\\
  \caption{\label{fig:keyspace-analysis_throughput}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-dynamic.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize-dynamic}}
\end{figure}
\begin{figure}[tb]
  \noindent\includegraphics[width=0.5\textwidth]{figures/keyspace-analysis_cachesize-caveats.png}\\
  \caption{\label{fig:keyspace-analysis_cachesize-caveats}}
\end{figure}