### 2017-11-26 paper/sections/5_file-systems.tex
paper: addresses #39
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Towards General Data Management Policies}
\label{sec:scope}

\begin{figure*}[t!]
    \centering
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \footnotesize
        \centering
        \begin{minted}[xleftmargin=3em,linenos]{lua}
function when()
  if server[whoami]['cachesize'] > n then
    if server[whoami]['cachesize'] > 100K then
      WRstate(1)
    end
    if RDstate() == 1 then
      return true
    end
  end
  return false
end
        \end{minted}
        \caption{ParSplice cache management policy that absorbs the burstiness of the
        workload before switching to a constrained cache.  The
        performance/utilization for different  \texttt{n} is in
        Figure~\ref{fig:methodology-tradeoff}. \label{src:lru-dyn}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.48\textwidth}
        \centering
        \footnotesize
        \begin{minted}[xleftmargin=3em,linenos]{lua}
local function when()
  if servers[whoami]["load"] > target then
    overloaded = RDstate() + 1
    WRstate(overloaded)
    if overloaded > 2 then
      return true
    end
  end
  else then WRstate(0) end
  return false
end
        \end{minted}
	\caption{CephFS file system metadata load balancer, designed in 2004
        in~\cite{weil:sc2004-dyn-metadata}, reimplemented in Lua
        in~\cite{sevilla:sc15-mantle}. This policy has many similarities to the
        ParSplice cache management policy.\label{src:lua-cephfs}}
    \end{subfigure}
    \caption{ParSplice's cache management policy has the same components as
    CephFS's load balancing policy.}
\end{figure*}

%\begin{itemize}
%  \item Lustre Trace
%  \item LinkedIn Trace
%  \item Nathan's Trace
%\end{itemize}

In the previous section, we used our data management language and the Mantle
policy engine to design effective cache management strategies for a new service
and domain. In this section, we compare and contrast the policies examined for
file system metadata load balancing in~\cite{sevilla:sc15-mantle} with the ones
we designed and evaluated above for cache management in ParSplice. The
similarities show how the ``when"/``where"/``how much" abstractions, data
management language, and policy engine may be widely applicable to other data
management techniques, such as QoS, scheduling, and batching.

\subsection{Using Load Balancing Policies for Cache Management}

%- memory utilization/heuristics
From a high-level the cache management policy we designed in
Figure~\ref{src:lru-dyn} trims the cache if the cache reaches a certain size
{\it and} if it has already absorbed the initial burstiness of the workload.
Much of this implementation was inspired by the file system metadata load
balancing policy in Figure~\ref{src:lua-cephfs}, which was presented
in~\cite{sevilla:sc15-mantle}. That policy migrates file system metadata if the
load is higher than the average load in the cluster {\it and} the current
server has been overloaded for more than two iterations. The two policies have
the following in common:

\textbf{Condition for ``Overloaded"} (Fig.~\ref{src:lru-dyn}: Line 2;
Fig.~\ref{src:lua-cephfs}: Line 2) - these lines detect whether the node is
overloaded using the load calculated in the load callback (not shown). While
the calculations and thresholds are different, the way the loads are used is
exactly the same; the ParSplice policy flags the node as overloaded if the
cache reaches a certain size while the CephFS policy compares the load to other
nodes in the system.

\textbf{State Persisted Across Decisions} (Fig.~\ref{src:lru-dyn}: Lines 4,6;
Fig~\ref{src:lua-cephfs}: Lines 3,4,9) - these lines use Mantle to write/read state
from previous decisions.  For ParSplice, we save a boolean that indicates
whether we have absorbed the workload's initial burstiness. For CephFS, we save
the number of consecutive instances that the server has been overloaded. We
also clear the count (Line 10) if the server is no longer overloaded.

\textbf{Multi-Policy Strategy} (Fig.~\ref{src:lru-dyn}: Line 6;
Fig.~\ref{src:lua-cephfs}: Line 5) - after determining that the node is
overloaded, these lines add an additional condition before the policy enters a
data management state.  ParSplice trims its cache once it eclipses the
``absorb" threshold while CephFS allows balancing when overloaded for more than
two iterations. The persistent state is essential for both of these
policy-switching conditions.

\subsection{Using Cache Management Policies for Load Balancing}

%- how many inodes to keep in mds/client caches
%- affects LB; increases capacity of single MDS (reduces req/mem pressure)
%- affects LB; tells us which part of the cache to keep local

% Load balancing in FSs
The abstractions in Mantle were designed for file system metadata load
balancing across a cluster of dedicated metadata servers, where spreading
requests across servers improves performance. But another technique to reduce
request load is caching. If clients and servers maintain a consistent cache,
the client can do operations locally without contacting the metadata server.
The caches in CephFS have the same performance and utilization trade-off as
ParSplice, where large caches improve performance at the expense of a higher
memory footprint. High memory utilization is problematic when a single metadata
server maintains consistent caches with many clients.  Using cache management
policies from the previous two sections has two benefits for load balancing:
(1) increases the capacity of a single metadata server, and (2) helps identify
which parts of the cache to keep local to a server. For example, applying the
key access pattern detection algorithm from
Section~\S\ref{sec:regime-detection} to a similar file system access pattern
has the potential to identify when to migrate keys and how many keys to
migrate.

% burstiness of creates then compile
To explore this example in more detail, consider the trace of metadata requests
for compiling code in CephFS shown in Figure~\ref{fig:compile-ops}.  That trace
shows the number of file system metadata requests serviced by the metadata
server when uncompressing (\texttt{untar}), compiling (\texttt{make}), and
deleting (\texttt{rm}) the source code for the Linux kernel. If the system
knows that the job phases would progress from many creates, to many lookups, to
many deletes, then it could size its caches accordingly. For example, the file
system could cache none of the metadata from the \texttt{untar} phase and run
key access pattern detection during the \texttt{make} phase, resulting in the metadata
server/clients only caching metadata that is repeatedly used. For the job in
Figure~\ref{fig:compile-ops}, this would fill up the cache to only 20K inodes
(the unit of metadata for file systems) instead of 60K, resulting in almost
40MB of memory savings (since an inode is about
1KB\footnote{http://docs.ceph.com/docs/master/dev/mds\_internals/data-structures/})
without sacrificing performance.

%- strategy: cache all inodes on create
\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/compile-ops.png}\\
\caption{File system metadata requests for a compile job show workload phases
characterized by a dominant request type ({\it e.g.}, creates for ``untar").
Using ParSplice cache management policies for these types of workloads would
reduce memory pressure without sacrificing performance.
\label{fig:compile-ops}}
\end{figure}

% other use cases caches/tiers 
Storage systems have many other data management techniques that would benefit
from the caching policies developed in Sections~\S\ref{sec:arch-specific}
and~\ref{sec:dom-specific}. For example, Ceph administrators can use the
policies in ParSplice to automatically size and manage cache
tiers\footnote{http://docs.ceph.com/docs/master/rados/operations/cache-tiering/},
caching on object storage devices, or in the distributed block
devices\footnote{http://docs.ceph.com/docs/master/rbd/rbd-config-ref/}.
Integration with Mantle would be straightforward as it is merged into Ceph's
mainline\footnote{http://docs.ceph.com/docs/master/cephfs/mantle/} and the
three caching subsystems mentioned above already maintain key access traces. We
hypothesize that since this is all software defined caching, something more
clever than LRU would improve cache utilization without sacrificing too much
performance.

\subsection{Other Use Cases}
% we said this is good for LB/Cache management... what about others??
% QoS: when to move clients, where to move clients, how much of the reservation to move
% Other cache management: MRU
% Scheduling: how much of a resource to allocate; more than just LRU (RAD)
%\subsection{Visualizing File System Traces like ParSplice Keyspace Traces}