### 2017-11-26 paper/sections/4_dynamic-policies.tex

%\subsection{Trajectory Length}
%
%\subsection{Cloud Techniques: Elastic Search}
%
%What if we re-provision resources in response to events outside the
%application's control, such as a slow Lustre.
%
%\subsection{Future work}
%
%How Much: cache policy from past, regime detection, Belady's Min

\section{Cache Management Using System Architecture Knowledge}
\label{sec:arch-specific}

Using the Mantle policy engine, we test a variety of cache management tools and
algorithms on the worker using the keyspace analysis in
Section~\S\ref{sec:parsplice-keyspace-analysis}.  The evaluation metric is the
accuracy and runtime of each strategy; the policy should be accurate enough so
as to sacrifice negligible performance and fast enough to run as often as we
want to detect key access patterns.  First we sized the cache according to our
system specific knowledge. By ``system", we mean the hardware and software of
the storage hierarchy. We look at request rate, unique keys in a sliding
window, and bandwidth capabilities. For example, we know that LevelDB cannot
handle high IO request rates.

% technical details
The results for different cache sizes for a growth rate of \(\Delta_1\) over a
2.5 hour run across 256 workers is shown in
Figure~\ref{fig:methodology-tradeoff}.  ``Baseline" is the performance of
unmodified ParSplice  measured in trajectory duration (\(y\) axis) and
utilization is measured with memory footprint of just the cache (\(y2\) axis).
The middle graph shares the \(y\) axis and shows the trade-off of using a basic
LRU-style cache for different cache sizes, where the penalty for a cache miss
is retrieving the data from the persistent database.  We evict keys (if
necessary) at every operation instead of when segments complete because the
cache fills up too quickly otherwise. This is implemented using a ``when"
policy of \texttt{server[whoami]['cachesize']>n} and a ``how much" policy of
\texttt{servers[whoami]['cachesize']-n}.  The error bars are the standard
deviation of 3 runs.  Although the keyspace grows to 150K, a 100K key cache
achieves 99\% of the performance. Decreasing the cache degrades performance and
predictability.

%\begin{figure}[h] \footnotesize \begin{minted}{lua}
%server[whoami]['cachesize'] > n \end{minted} \end{figure}

%and ``how much":
%\begin{figure}[h]
%\footnotesize
%\begin{minted}{lua}
%servers[whoami]['cachesize'] - n
%\end{minted}
%\end{figure}

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/methodology-tradeoff.png}\\
\caption{The performance and resource utilization trade-off for different cache
management policies. ``None" is ParSplice unmodified, ``Fixed Sized Cache"
limits the size of the cache by evicting the least recently used items, and
``Multi-Policy Cache" switches to a fixed sized cache after absorbing the
initial burstiness of the workload.  The size of the fixed sized cache for each
experiment is on the \(x\) axis.  \label{fig:methodology-tradeoff}}
\end{figure}
%
%\footnotesize \centering \begin{minted}[xleftmargin=3em,linenos]{lua} function
%when() if server[whoami]['cachesize'] > n then return true end return false
%end
%
%function howmuch() if servers[whoami]['cachesize'] > n return
%servers[whoami]['cachesize'] - n end return 0 end \end{minted} \caption{Policy
%that implements a basic LRU cache. The performance and memory utilization for
%different values of \(n\) are graphed in
%Figure~\ref{fig:methodology-tradeoff}.  \label{src:lru}} \end{figure}

% why is the performance lower for smaller caches?
But the top graph in Figure~\ref{fig:motivation-regimes} suggests that a
smaller cache size should suffice, as only 100 keys seem to be active at any
one time.  It turns out that the unique keys plotted in
Figure~\ref{fig:motivation-regimes} are per second and are not representative
of the actual active keyspace; the number of active keys is larger than 100, as
some keys may be accessed at time \(t_0\), not in \(t_1\), and then again in
\(t_2\). Because the cache is too small, reads and writes hit the persistent
database, where the excessive traffic triggers a LevelDB compaction and reads
block.  To avoid overloading the storage hierarchy with requests, we design a
multi-part policy that switches between:

\begin{itemize}
  \item unlimited growth policy: cache increases on every write
  \item \(n\) key limit policy: cache constrained to \(n\) keys
\end{itemize}

% results: same level of performance can be achieved 
The key observation is that small caches incur too much load on the persistent
database at the beginning of the run but should suffice after the initial read
flash crowd passes because the keyspace is far less active.  We program Mantle
to trigger the policy switch at 100K keys to absorb the flash crowd at the
beginning of the run. Once triggered, keys are evicted to bring the size of the
cache down to the threshold.  The actual policy is shown and described in more
detail in the next section in Figure~\ref{src:lru-dyn}.  The plot on the right
side of Figure~\ref{fig:methodology-tradeoff} shows the performance/utilization
trade-off of the multi-part policy, where the cache sizes for the \(n\) key
limit policy are along the \(x\) axis.  They show better performance than the
single \(n\) key policies. The performance and memory utilization for a 100K
key cache size is the same as the 100K bar in the ``Multi-Part Policy" graph in
Figure~\ref{fig:methodology-tradeoff} but the rest reduce the size of the
keyspace after the read flash crowd.  We see the worst performance when the
engine switches to the 10 key limit policy, which achieves 94\% of the
performance while only using 40KB of memory.

% caveats: it is calculating 90% of the trajectory, memory value reported is final
\subsubsection*{Caveats}

The results in Figure~\ref{fig:methodology-tradeoff} are slightly deceiving for
two reasons: (1) segments take longer to generate later in the run and (2) the
memory footprint is the value at the end of 2.5 hours.  For (1), the trajectory
length vs.  wall-clock time curves down over time; as the nanoparticle grows
it takes longer to generate segments so by the time we reach 2 hours, over 90\%
of the trajectory is already generated.  For (2), the memory footprint rises
until we reach 100K key switch threshold at 0.4GB, as shown by the
``\(\Delta_1\), Multi-Policy" curve in Figure~\ref{fig:memory-vs-time} but in
Figure~\ref{fig:methodology-tradeoff} we plot the final value. As a result, we
can only conclude that this policy only works well for this 2.5 hour run.

% but the result is still valid
Despite these caveats, the result is still valid: we found a multi-policy cache 
management strategy that absorbs the cost of a high read throughput on a small
keyspace and reduces the memory pressure for a 2.5 hour run. Our experiments
show the effectiveness of the policy engine we integrated into
ParSplice, not that we were able to identify the best policy for all system
setups ({\it i.e.} different ParSplice parameters, number of worker tasks, and
job lengths).  To solve that problem, we need a way to identify what thresholds
we should use for different job permutations.

\section{Cache Management Using Domain-Specific Application Knowledge}
\label{sec:dom-specific}

\begin{figure*}[t!]
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
	\includegraphics[width=0.89\textwidth]{figures/keyspace-zoomed.png}\\
\caption{Key activity for a 4 hour run shows groups of accesses to the same
subset of keys. Detecting these access patterns leads to a more accurate cache
management strategy, which is discussed in
Section~\S\ref{sec:regime-detection}.\label{fig:keyspace-zoomed}}
    \end{subfigure}%
    ~ 
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[width=1\textwidth]{figures/memory-vs-time.png}\\
	\caption{Memory utilization for ``No Cache Management" (unlimited cache
growth), ``Multi-Policy" (absorbs initial burstiness of workload) , and
``Dynamic Policy" (sizes cache according to key access patterns).
\label{fig:memory-vs-time}}
    \end{subfigure}%
    ~
    \begin{subfigure}[t]{0.32\textwidth}
        \centering
        \includegraphics[width=0.86\textwidth]{figures/dscache-vs-none.png}\\
	\caption{Performance/Utilization for dynamically sized cache (DSCache)
policy. With negligible performance degradation, DSCache adjusts to different
\(\Delta\)s over 2\(\times\) memory in the best case.
\label{fig:dscache-vs-none}}
    \end{subfigure}%
    \caption{Different cache management policies tested over the Mantle policy engine.}
\end{figure*}

%\begin{figure}[t]
%  \noindent\includegraphics[height=5cm,width=0.4\textwidth]{figures/keyspace-zoomed.png}\\
%
%  \caption{Key activity for a 4 hour run shows groups of accesses to the same
%  subset of keys. Detecting these access patterns leads to a more accurate cache
%  management strategy.\label{fig:keyspace-zoomed}}
%
%\end{figure}
%
%\begin{figure}[t]
%\noindent\includegraphics[width=0.5\textwidth]{figures/dscache-vs-none.png}\\
%
%\caption{The performance and resource utilization trade-off for the dynamically
%sized cache (DSCache) policy compared to no cache management. With a negligible
%performance impact, DSCache adjusts to different initial conditions
%(\(\Delta\)) automatically and saves over 2\(\times\) memory in the best case.
%\label{fig:dscache-vs-none}}
%\end{figure}
%
%\begin{figure}[t]
%\noindent\includegraphics[width=0.5\textwidth]{figures/memory-vs-time.png}\\
%
%\caption{Memory utilization of different cache management policies.  ``No Cache
%Management" allows unlimited cache growth, ``Multi-Policy" absorbs the initial
%burstiness of the workload before switching to a more constrained cache, and
%the ``Dynamic Policy" curves size the cache according to key access patterns.
%We show a \(\Delta_2\) growth rate to demonstrate the dynamic policy's ability
%to adjust to a different set of initial conditions.\label{fig:memory-vs-time}}
%\end{figure}

Feeding domain-specific knowledge about the ParSplice application into a policy
leads to more a accurate cache management strategy.  The goal of the following
sections is not to find an optimal solution, as this can be done with parameter
sweeps for thresholds; rather, we try to find techniques that work for a range
of inputs and system setups.

Figure~\ref{fig:keyspace-zoomed} shows which keys (\(y\) axis) are accessed by
tasks over time (\(x\) axis). The groups of accesses to a subset of keys occurs
because molecules are stuck in deep trajectories. Recall that the cache stores
the molecules' EOM minima, which is the smallest effective energy that a
molecule observes during its trajectory. So molecules stuck in deep
trajectories explore the same minima until they can escape to a new set of
states. This exploration of the same set of states is called a superbasin.  In
Figure~\ref{fig:keyspace-zoomed}, superbasins are never re-visited because the
simulation only adds molecules; we can never reach a state with less molecules.
This is why keys are never re-accessed.  Despite these patterns, the following
characteristics of superbasins make it hard to detect them:

Detecting these superbasins can lead to more effective cache management
strategies because the height of the groups of key accesses is ``how much" of
the cache to evict and the width of the groups of key accesses is ``when" to
evict values from the cache.  The zoomed portion of
Figure~\ref{fig:keyspace-zoomed} shows how a single superbasin affects the key
accesses. Moving along the \(x\) axis shows that the number of unique keys
accessed over time grows while moving along the \(y\) axis shows that early
keys are accessed more often.

\begin{itemize}

  \item superbasin key accesses are random and there is no threshold ``minimum distance
  between key access" that indicates we have moved on to a new superbasin

  \item superbasins change immediately

  \item the number of keys a superbasin accesses differs from other superbasins

\end{itemize}

Below we describe the policies we implemented and deployed using Mantle.

\subsection{Failed Strategies}

To detect the access patterns in Figure~\ref{fig:keyspace-zoomed}, we try a
variety of techniques using Mantle. Unfortunately, we found that the following
techniques proliferate more parameters that need to be tuned per
hardware/software configuration. Furthermore, we find that many of the metrics
do not signal a new set of key accesses.  Below, we indicate with quotes which
parameters we need to add for each technique with and the value we find to work
best, via tuning and parameter sweeps, for one set of initial conditions.

\begin{itemize}

  \item Statistics: decay on each key counts down until 0; 0-valued keys are
evicted.  ``history-of-key-accesses", set to 10 seconds, to evict keys.

  \item Calculus: use derivative to strip away magnitudes; use large positive
slopes followed by large negative slope as signal for new set of key accesses.
``Zero-crossing", set to 40 seconds, for distance between small/large spikes to
avoid false positives; `window size", set to 200 seconds, for the size of the
moving average

  \item K-Means Clustering fails because ``K" is not known {\it a-priori} and
groups of key accesses are different size. ``K", set to 4, for the number of clusters in the data
using the sum of the distances to the centroid

  \item DBScan: finds clusters using density as a metric. ``Eps", set to 20, for
max distance between 2 samples in same neighborhood; ``Min", set to 5, for the
number of samples per core

  \item Anomaly Detection: size of the image is too big and bottom edges are
not thick enough

\end{itemize}

\subsection{Dynamically Sized Cache: Access Pattern Detection}
\label{sec:regime-detection}

After trying these techniques we found that the basic algorithm in
Figure~\ref{src:dyn-cache} works best. The algorithm detects groups of key
access patterns, which we call ``fans", by iterating backwards through the key
access trace, finding the lowest key ID, and comparing against the lowest key
ID we have seen so far (Line 7). We also maintain the top and bottom of each
group of key accesses (Line 13) so we can tell the ``how much" policy the
number of keys to evict (Line 19).  The algorithm iterates backwards over the
key access trace because a change in the minimum value signals a new group of
key accesses. No signal exists iterating left to right, as the maximum value
always increases and the minimum values at the bottom of each group of key
accesses are sparse.  For example, the maximum distance between values along
the bottom edge of the zoomed group of key accesses in
Figure~\ref{fig:keyspace-zoomed} is 125 seconds, while the maximum distance
between minimum values for the group of key accesses before is 0 seconds. As a
result of this sparseness, iterating left to right requires a ``window size"
parameter to determine when we think a minimum value will not show up again.
We exchange runtime, in the form of an O(\(n\)) algorithm where \(n\) is the
number events, for simplicity because this approach avoids adding new
thresholds for key access pattern detection ({\it e.g.}, space between key
accesses, space between key IDs, and window size of consecutive key accesses).

The performance and memory utilization is shown by the ``DSCache" bars in
Figure~\ref{fig:dscache-vs-none}. Without sacrificing performance (trajectory
length), the dynamically sized cache policy uses between 32\%-66\% less memory
than the default ParSplice configuration (no cache management) for the 3
initial conditions we test. The memory usage is shown by the ``Dynamic Policy"
curves in Figure~\ref{fig:memory-vs-time}, where the behavior resembles the key
access patterns in Figure~\ref{fig:keyspace-zoomed}\footnote{The memory usage
is not {\it exactly} the same because these are two different runs;
Figure~\ref{fig:keyspace-zoomed} has key activity tracing turned on, which
reduces performance.}.  We also show a \(\Delta_2\) growth rate to demonstrate
the dynamic policy's ability to adjust to a different set of initial
conditions.

\begin{figure}[h]
\footnotesize
\begin{minted}[xleftmargin=3em,linenos]{lua}
d = timeseries()          -- provided by mantle
ts, id = d:get(d:size())  -- provided by mantle
fan  = {start=nil, finish=ts, top=0, bottom=id}
fans = {}
for i=d:size(),1,-1 do    -- detect all fans
  ts, id = d:get(i)
  if id < fan['bottom'] then
    fan['start']  = ts
    fans[#fans+1] = fan 
    fan = {start=nil, finish=ts, top=0, bottom=id}
  end

  if id > fan['top'] then fan['top'] = id end 
end
fan['start'] = 0 
fans[#fans+1] = fan

if #fans < 2 then return false else
  WRstate(fans[#fans-1]['top']-fans[1]['bottom'])
  return true
end
\end{minted}
\caption{The dynamically sized cache policy, whose performance is shown in 
Figure~\ref{fig:cache-management}, iterates backwards over timestamp-key pairs
and detects when key accesses move on to a new subset of
keys. We refer to groups of key accesses as ``fans".\label{src:dyn-cache}}
\end{figure}