### 2017-11-26 paper/sections/6_conclusion.tex

\section{Related Work}

Key-value storage organizations for scientific applications is a field gaining
rapid interest. In particular, the analysis of the ParSplice keyspace and the
development of an appropriate scheme for load balancing is a direct response to
a case study for computation caching in scientific
applications~\cite{jenkins:ipdsw17-mochi}. In that work the authors motivated
the need for a flexible load balancing \emph{microservice} to efficiently scale
a memoization microservice. Our work is also heavily influenced by the
Malacology project~\cite{sevilla:eurosys17-malacology} which seeks to provide
fundamental services from within the storage system ({\it e.g.}, consensus) to
the application.  Our plan is to use
MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end key-value store
because it was designed for HPC and has the proper mechanisms for migration
already implemented.

State-of-the-art distributed file systems partition write-heavy workloads and
replicate read-heavy workloads, similar to the approach we are advocating
here.  IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients
write to different partitions by grabbing leases and caching ancestor metadata
for path traversal. ShardFS takes the replication approach to the extreme by
copying all directory state to all nodes. The Ceph file system
(CephFS)~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
techniques to a lesser extent; directories can be replicated or sharded but the
caching and replication policies are controlled with tunable parameters.  These
systems still need to be tuned by hand with {\it ad-hoc} policies designed for
specific applications.  Setting policies for migrations is arguably more
difficult than adding the migration mechanisms themselves.  For example,
IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
partitioning directories at a \emph{predefined} threshold. Mantle makes headway
in this space by providing a framework for exploring these policies, but does
not attempt anything more sophisticated ({\it e.g.}, machine learning) to create
these policies.

% ml and autotuning
Auto-tuning is a well-known technique used in
HPC~\cite{behzad:sc2013-autotuning, behzad:techreport2014-io-autotuning}, big
data systems systems~\cite{herodotou_starfish_2011}, and
databases~\cite{schnaitter_index_2009}.  Like our work, these systems focus on
the physical design of the storage ({\it e.g.} cache size) but since we focused
on a relatively small set of parameters (cache size, migration thresholds), we
did not need anything as sophisticated as the genetic algorithm used
in~\cite{behzad:sc2013-autotuning}.  We cannot drop these techniques into
ParSplice because the magnitude and speed of the workload hotspots/flash crowds
makes existing approaches less applicable.

\section{Future Work}

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across domains
and services. The value of such a collection eases the burden of policy
development and paves the way for solutions that remove the administrator from
the development cycle, such as (1) adaptable policies that automatically switch
to new strategies when the current strategy behaves poorly ({\it e.g.}, thrashing,
making no progress, etc.), and (2) policy generation, where new policies are
constructed automatically by examining the collection of existing policies.
Ultimately, we hope that this automation enables control of policies by
machines instead of administrators.

\section{Conclusion}

Data management encompasses a wide range of techniques that vary by domain and
service. Yet, the techniques require policies that shape the decision making
and finding the best policies is a difficult, multi-dimensional problem.  We
iterate to a custom solution for our target application that uses workload
access patterns to size its caches. Without tuning or parameter sweeps, our
solution saves memory without sacrificing performance for a variety of initial
conditions, including the scale, duration, configuration, and hardware of the
simulation. More importantly, rather than attempting to construct a single,
complex policy that works for a variety of scenarios, we instead use the Mantle
framework to enable software-defined storage systems to flexibly change
policies as the workload changes.  We also observe that many of the primitives
and strategies have enough in common with data management in file systems that
they both can be expressed with similar semantics and a general policy engine.
Our prototype lays the groundwork for future work in adaptable policies and
policy generation, which ushers in an era where policies are controlled by
machines instead of administrators.