### 2017-11-27 paper/sections/3_methodology.tex
paper: addresses #53
Signed-off-by: Michael Sevilla <mikesevilla3@gmail.com>

\section{Methodology}

%  \item background: CephFS implementation
To explore software-defined cache management, we use the data management
language and policy engine presented in~\cite{sevilla:sc15-mantle}. The
prototype in that paper was called Mantle and it was originally built on CephFS
so administrators could control file system metadata load balancing policies.
We now refer to Mantle as a policy engine that supports our data management
language.  The basic premise is that data management policies can be expressed
with a simple API consisting of ``when", ``where", and ``how much". The ``when" policy
controls how aggressive or conservative the decisions are; ``where" controls
how distributed or concentrated the data should be; and ``how much" controls
the amount of data that should be sent. There is also a ``load" policy that
lets administrators specify how to collapse many metrics into a single load
metric ({\it e.g.}, \(2\times\texttt{cpu} + 3\times\texttt{memory usage}\)).

The succinctness of the API lets users inject multiple, possibly dynamic, policies. In
this work we focus on a single node, so the ``where" policy is not used. When
we move ParSplice to a distributed key-value store back-end, the ``where" policy
will be used determine which key-value pairs should be moved to which node.

\subsection{Extracting Mantle as a Library}
\label{sec:extracting}

We extracted Mantle as a library and Figure~\ref{fig:mantle} shows how it is
linked into a storage system.  Administrators write policies in Lua from whatever
domain they choose ({\it e.g.}, statistics, machine learning, storage system)
and the policies are embedded into the runtime by Mantle.  We chose Lua for
simplicity, performance, and portability; it is a scripting language with
simple syntax, which allows administrators to focus on the policies themselves;
it was designed as an embeddable language, so it is lightweight and does less
type checking; and it interfaces nicely with C/C++.  When the storage system makes
decisions it executes the administrator-defined policies for when/where/how
much and returns a decision.  To do this, the storage system needs to be modified to
(1) provide an environment of metrics and (2) identify where policies are set.
These modification points are shown by the colored boxes in
Figure~\ref{fig:mantle} and described below.

\begin{figure}[t]
  \noindent\includegraphics[width=0.5\textwidth]{figures/mantle.png}\\

  \caption{Extracting Mantle as library.\label{fig:mantle}}

\end{figure}

\begin{table}
  \centering
  \begin{tabular}{ r | l | l }
  Metrics     & Data Structure & Description \\\hline
  Cluster     & \{server \(\rightarrow\) \{metric \(\rightarrow\) val\}\}
              & resource util. for servers \\
  Time Series & [(ts, val), ..., (ts, val)]
              & accesses by timestamp (ts) \\
  && \\
              & Storage System & Example \\\hline
  Cluster     & File Systems & CPU util., Inode reads \\
              & ParSplice    & CPU util., Cache size \\
  Time Series & File Systems & Accesses to directory \\
              & ParSplice    & Accesses to key in DB\\
  \end{tabular}
  \caption{Types of metrics exposed by the storage system to the policy engine using Mantle.\label{table:metrics}}
\end{table}

\subsubsection{Environment of Metrics} storage systems expose \textbf{cluster} metrics
for describing resource utilization and \textbf{time series} metrics for
describing accesses to some data structure over time. Table~\ref{table:metrics}
shows how these metrics are accessed from the policies written by
administrators.

For cluster metrics, the storage system passes a dictionary to Mantle. Policies access
the cluster metric values by indexing into a Lua table using \texttt{server}
and \texttt{metric}, where server is a node identifier ({\it e.g.}, MPI Rank,
metadata server name) and metric is a resource name.  Metrics used for file
system metadata load balancing are shown by the ``environment" box in
Figure~\ref{fig:mantle}. The measurements and exchange of metrics between
servers is done by the storage system; Mantle in CephFS leverages metrics from other
servers collected using CephFS's heartbeats.  For example, a policy written for
an MPI-based storage system can access the CPU utilization of the first rank in a
communication group using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
load = servers[0]['cpu']
\end{minted}
\end{figure}

For time series metrics, the storage system passes an array of \texttt{(timestamp,
value)} pairs to Mantle and the policies can iterate over the accesses.  The
storage system uses a pointer to the time series to facilitate time series with many
values, like accesses to a database or directory in the file system namespace.
This decision limits the time series metrics to only include values from the
{\it current } node, although this is not a limitation of Mantle itself.  For
example, a policy that uses accesses to a directory in a file system as a
metric for load collects that information using:

\begin{figure}[h]
\footnotesize
\begin{minted}{lua}
d = timeseries()       -- data (d) from storage system
for i=1,d:size() do    -- iterate over timeseries
  ts, value = d:get(i) -- index into timeseries 
  if value == 'mydirectory' then
    count = count + 1
  end
end
\end{minted}
\end{figure}

% policies themselves
\subsubsection{Policies Written as Callbacks} the ``callback" box in
Figure~\ref{fig:mantle} shows an example policy for ``when()", where the
current server (\texttt{whoami}) migrates load if it is has load
(\texttt{>0.1}) and if its neighbor server (\texttt{whoami + 1}) does not have
load (\texttt{<0.1}). The load is calculated using the metrics provided by the
environment.

Mantle also provides functions for persisting state across decisions.
\texttt{WRState(s)} saves state \texttt{s}, which can be a number or boolean
value, and \texttt{RDState()} returns the state saved by a previous iteration.
For example, a ``when" policy can avoid trimming a cache or migrating data if
it had performed that operation in the previous decision.

%\begin{figure}[h]
%\footnotesize
%\begin{minted}{lua}
%if RDState() == 1 then
%  WRState(0)   -- the next decision will return true
%  return false
%else then
%  WRState(1)   -- the next decision will return false
%  return true
%end
%\end{minted}
%\end{figure}

%of timestamp (ts), value pairs representing accesses over time types of
\subsection{Integrating Mantle into ParSplice}

% env of metrics
Using Mantle cluster metrics, we expose cache size, CPU utilization, and memory
pressure of the worker node to the cache management policies. In
Section~\S\ref{sec:arch-specific} we only end up using the cache size although
the other metrics proved to be valuable debugging tools. Using Mantle time
series metrics, we expose accesses to the cache as a list of \texttt{timestamp,
key} pairs. In Section~\S\ref{sec:dom-specific}, we explore a key access
pattern detection algorithm that uses this metric.

% where policies are made
We link Mantle directly into the worker cache circled in red in
Figure~\ref{fig:parsplice}. We put the ``when" and ``how much" callbacks
alongside code that checks for memory pressure. It is executed right before the
worker processes incoming and outgoing put/get transactions to the cache. We
only do cache management once every second to avoid maintaining the cache for
every request. We expected to have to increase this polling interval to
accommodate more complex policies but even our most complicated policy in
Section~\S\ref{sec:dom-specific} had a negligible effect on performance when
executed every second (within the standard deviation for multiple runs when
compared against a policy that returns immediately). This may be because the
worker is not overloaded and the bottleneck is somewhere else in the system.
As stated previously, we do not use the ``where" part of Mantle because we
focus on a single node, but this part of the API will be used when we move the
caches and storage nodes to a key-values store back-end that uses key load
balancing and repartitioning.