### 2017-11-28 paper/sections/1_introduction.tex
Changed paragraph before list of contribs

\begin{abstract}

Our analysis of the key-value activity generated by the ParSplice molecular
dynamics simulation demonstrates the need for more complex cache management
strategies. Baseline measurements show clear key access patterns and hot spots
that offer significant opportunity for optimization. We use the data management
language and policy engine from the Mantle system to dynamically explore a
variety of techniques, ranging from basic algorithms and heuristics to
statistical models, calculus, and machine learning. While Mantle was originally
designed for distributed file systems, we show how the collection of
abstractions effectively decomposes the problem into manageable policies for a
different application and storage system.  Our exploration of this space results in a
dynamically sized cache policy that sacrifices negligible performance while
using 32-66\% less memory than the default ParSplice configuration.

\end{abstract}

\section{Introduction}

Storage systems use software-based caches to improve performance but the
policies that guide what data to evict and when to evict vary with the use
case. For example, caching file system metadata on clients and servers reduces
the number of remote procedure calls and improves the performance of
create-heavy workloads common in HPC~\cite{ren:sc2014-indexfs,
patil:fast2011-giga+}. But the policies for what data to evict and when to
evict are specific to the application's behavior and the system's configuration
(hardware, settings, etc.) so a new workload may prove to be a poor match for
the selected caching
policy~\cite{xiao:socc15-shardfs,brandt:msst2003-lh,sevilla:sc15-mantle,
weil:sc2004-dyn-metadata, weil:osdi2006-ceph}. In this paper, we evaluate a
variety of caching policies using our data management language/policy engine and
arrive at a customized policy that works well for our target application.  This
process of trying general policies and quickly iterating to a customized
solution shows that our prototype can be adapted to different workloads.

Our example application, ParSplice~\cite{perez:jctc20150parsplice}, is
representative of an important class applications with similar working set
behaviors that extensively use software-based caches.  ParSplice is a molecular
dynamics simulation that uses a hierarchy of caches and a single persistent
key-value store to store both observed minima across a molecule's equation of
motion (EOM) and the hundreds or thousands of partial trajectories calculated
each second during a parallel job.  This workload is pervasive across
simulations that rely on a mesh-based decomposition of a physical region and
result in millions or billions of mesh cells, where each cell contains
materials, pressures, temperatures and other characteristics that are required
to accurately simulate phenomena of interest.  The fine-grained data annotation
capabilities provided by key-value storage is a natural match for these types
of scientific simulations.  Unfortunately, simulations of this size saturate
the capacity and bandwidth capabilities of a single node so we need more
effective data management techniques.

\begin{figure}[t]
\noindent\includegraphics[width=0.5\textwidth]{figures/cache-management.png}\\
\caption{Using our data management language and policy engine, we design a
dynamically sized caching policy (thick line) for ParSplice.  Compared to
existing configurations (thin lines with \(\times\)'s), our solution saves the most
memory without sacrificing performance and works for a variety of inputs.
\label{fig:cache-management}}
\end{figure}

The biggest challenge for ParSplice is properly sizing the caches in the
storage hierarchy.  The memory usage for a single cache that stores molecule
coordinates is shown in Figure~\ref{fig:cache-management}, where the thin solid
lines marked with \(\times\)'s are the existing configurations in ParSplice.
The default configuration uses an unlimited sized cache, shown by the ``No
Cache Management" line, but using this much memory for one cache is
unacceptable for HPC environments, where a common goal is to keep memory for
such data structures below 3\%\footnote{Anecdotally, we find this threshold
works well for HPC applications.  For reference, a 1GB cache for a distributed
file system is too large in LANL deployments.}. Furthermore, ParSplice deploys
a cache per 300 worker processes, so large simulations need more caches and
will use even more memory.  Users can configure ParSplice to evict data when
the cache reaches a threshold but this solution requires tuning and parameter
sweeps; the ``Cache (too small)" curve in Figure~\ref{fig:cache-management}
shows how a poorly configured cache can save memory but at the cost of
performance, which is shown by the text annotation to the right.  Even worse,
this threshold changes with different initial configurations and cluster setups
so tuning needs to be done for all system permutations.  Our dynamically sized
cache, shown by the thick line in Figure~\ref{fig:cache-management}, detects
key access patterns and re-sizes the cache accordingly.  Without tuning or
parameter sweeps, our solution saves more memory than a hand-tuned cache with a
negligible performance degradation (within the baseline's standard deviation),
works for a variety of initial conditions, and could generalize to similar
applications.

% What is Mantle
In this paper we are presenting the successful use of the Mantle policy language and engine to manage the dynamically-sized cache. Mantle provides a control plane that
injects policies into a running storage system, such as a file system or key-value store. While Mantle
was originally designed for file system metadata load balancing~\cite{sevilla:sc15-mantle},
we found that it works surprisingly well for specifying cache management policies without requiring users to possess extensive knowledge about the internals of storage systems. We show that our framework:

\begin{itemize}

  \item decomposes cache management into independent policies that can be
  dynamically changed, making the problem more manageable and easier to reason
  about.

  \item can deploy a variety of cache management strategies ranging from basic
  algorithms and heuristics to statistical models and machine learning.

  \item has useful primitives that, while designed for file system metadata
  load balancing, turn out to also be effective for cache management.

\end{itemize}

% this gives us many policies that are effective across disciplines
% - reuse: eases burden of writing policies
% - autonomic: lays groundwork for an adaptable policy that mixes/matches policies
% FUTURE WORK

This last contribution is explored in Sections~\S\ref{sec:arch-specific}
and~\S\ref{sec:dom-specific}, where we try a range of policies from different
disciplines; but more importantly, in Section~\S\ref{sec:scope}, we conclude
that the collection of policies we design for cache management in ParSplice
are very similar to the policies used to load balance metadata in the Ceph file
system (CephFS) suggesting that there is potential for automatically adapting
and generating policies dynamically.

%Manageable: abstracts away complexities of the system (pass around to others,
%use different strategies)