### 2017-11-28 paper/sections/6_conclusion.tex

\section{Related Work}

Key-value storage organizations for scientific applications is a field gaining
rapid interest. In particular, the analysis of the ParSplice keyspace and the
development of an appropriate scheme for load balancing is a direct response to
a case study for computation caching in scientific
applications~\cite{jenkins:ipdsw17-mochi}. In that work the authors motivated
the need for a flexible load balancing \emph{microservice} to efficiently scale
a memoization microservice. Our work is also heavily influenced by the
Malacology project~\cite{sevilla:eurosys17-malacology} which seeks to provide
fundamental services from within the storage system ({\it e.g.}, consensus) to
the application.  Our plan is to use
MDHIM~\cite{greenberg:hotstorage2015-mdhim} as our back-end key-value store
because it was designed for HPC and has the proper mechanisms for migration
already implemented.

State-of-the-art distributed file systems partition write-heavy workloads and
replicate read-heavy workloads, similar to the approach we are advocating here.
IndexFS~\cite{ren:sc2014-indexfs} partitions directories and clients write to
different partitions by grabbing leases and caching ancestor metadata for path
traversal. ShardFS~\cite{xiao:socc15-shardfs} takes the replication approach to
the extreme by copying all directory state to all nodes.
CephFS~\cite{weil:sc2004-dyn-metadata, weil:osdi2006-ceph} employs both
techniques to a lesser extent; directories can be replicated or sharded but the
caching and replication policies are controlled with tunable parameters.  These
systems still need to be tuned by hand with {\it ad-hoc} policies designed for
specific applications.  Setting policies for migrations is arguably more
difficult than adding the migration mechanisms themselves.  For example,
IndexFS/CephFS use the GIGA+~\cite{patil:fast2011-giga} technique for
partitioning directories at a \emph{predefined} threshold. Mantle makes headway
in this space by providing a framework for exploring these policies, but does
not attempt anything more sophisticated ({\it e.g.}, machine learning) to
create these policies.

% ml and autotuning
Auto-tuning is a well-known technique used in
HPC~\cite{behzad:sc2013-autotuning, behzad:techreport2014-io-autotuning}, big
data systems systems~\cite{herodotou_starfish_2011}, and
databases~\cite{schnaitter_index_2009}.  Like our work, these systems focus on
the physical design of the storage ({\it e.g.} cache size) but since we focused
on a relatively small set of parameters (cache size, migration thresholds), we
did not need anything as sophisticated as the genetic algorithm used
in~\cite{behzad:sc2013-autotuning}.  We cannot drop these techniques into
ParSplice because the magnitude and speed of the workload hotspots/flash crowds
makes existing approaches less applicable.

\section{Conclusion}

Data management encompasses a wide range of techniques that vary by application
and storage system.  Yet, the techniques require policies that shape the
decision making and finding the best policies is a difficult, multi-dimensional
problem.  We iterate to a custom solution for our target application that uses
workload access patterns to size its caches. Without tuning or parameter
sweeps, our solution saves memory without sacrificing performance for a variety
of initial conditions, including the scale, duration, configuration, and
hardware of the simulation. More importantly, rather than attempting to
construct a single, complex policy that works for a variety of scenarios, we
instead use the Mantle framework to enable software-defined storage systems to
flexibly change policies as the workload changes.  We also observe that many of
the primitives and strategies have enough in common with data management in
file systems that they both can be expressed with similar semantics.

This lays the foundation for future work, where we will focus on formalizing a
collection of general data management policies that can be used across
applications and storage systems.  The value of such a collection eases the
burden of policy development and paves the way for automated solutions such as
(1) adaptable policies that switch to new strategies when the current strategy
behaves poorly ({\it e.g.}, thrashing, making no progress, etc.), and (2)
policy generation, where new policies are constructed by examining the
collection of existing policies.  Ultimately, we hope that this automation
enables control of policies by machines instead of administrators.